---
title: "Mass Spectrometry-Based Single-Cell Proteomics Data Analysis"
author: "Zhicheng Yang"
date: "2026-04-22"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

prepare_child_rmd <- function(path) {
  lines <- readLines(path, warn = FALSE, encoding = "UTF-8")
  yaml_start <- which(trimws(lines) == "---")[1]

  if (is.na(yaml_start)) {
    stop(sprintf("No YAML header found in %s", path))
  }

  lines <- lines[yaml_start:length(lines)]
  yaml_marks <- which(trimws(lines) == "---")

  if (length(yaml_marks) >= 2) {
    body <- lines[(yaml_marks[2] + 1):length(lines)]
  } else {
    body <- lines
  }

  tmp <- tempfile(pattern = "child_", fileext = ".Rmd")
  writeLines(body, tmp, useBytes = TRUE)
  tmp
}

figure2_child <- prepare_child_rmd("Figure2.Rmd")
figure3_child <- prepare_child_rmd("Figure3.Rmd")
figure3_updated_child <- prepare_child_rmd("Figure3_20251125.Rmd")
figure4_child <- prepare_child_rmd("Figure4.Rmd")
```

# Overview

This notebook consolidates the analysis code from the original figure-specific R Markdown files into a single entry point.

The source files are kept alongside this notebook:

- `Figure2.Rmd`
- `Figure3.Rmd`
- `Figure3_20251125.Rmd`
- `Figure4.Rmd`

The original data directories and generated plot outputs are not bundled in this export repository.

# Figure 2

```{r figure2, child = figure2_child}
```

# Figure 3

## Original Version

```{r figure3, child = figure3_child}
```

## Updated Version (2025-11-25)

```{r figure3-updated, child = figure3_updated_child}
```

# Figure 4

```{r figure4, child = figure4_child}
```
