df_med <- df_log |>
  rowwise() |>
  mutate(
    one_med = median(c_across(starts_with("one_")), na.rm = TRUE),
    two_med = median(c_across(starts_with("two_")), na.rm = TRUE)
  ) |>
  ungroup() |>
  filter(!is.na(one_med), !is.na(two_med))
---
title: "Mass Spectrometry-Based Single-Cell Proteomics Data Analysis"
author: "Zhicheng Yang"
date: "2025-07-15"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    latex_engine: xelatex
    code_folding: hide
fontsize: 18pt
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.height = 6,
  fig.width = 8,render = F
)
setwd(".")
source("../../20240703function_combine.R")
library(Seurat)
library(scp)
library(ggplot2)
library(dplyr)
library(patchwork)
library(data.table)
library(QFeatures)
library(viridis)
library(stringr)
library(tidyr)
library(tibble)
library(SingleCellExperiment)
group_valuess = wheel("steelblue", num = 4)
```

```{r function}

spectronaut_to_iden<-function(data){
  data1<-data[,grep(".Quantity",colnames(data))]
  colnames(data1)<-unname(sapply(colnames(data1),function(x) strsplit(x,"\\.")[[1]][4]))
  data1[data1=="NaN"]<-NA
   p_num<-data.frame(colSums ( !is.na ( data1 )))
  colnames(p_num)<-"p_num"
  p_num$sample<-rownames(p_num)
  return(p_num)
}
add_meta <- function(df, method, metric, chars = "SPD"){
  pattern <- paste0("(?i)(\\d+)(?=\\s*", chars, ")")
  
  df %>%
    mutate(
      method = method,
      metric = metric,
      chars_num = coalesce(
        suppressWarnings(as.integer(str_extract(sample, pattern))),
        suppressWarnings(as.integer(str_extract(sample, "\\d+")))
      ),
      chars = factor(paste0(chars_num, " ", chars),
                     levels = unique(paste0(sort(unique(chars_num)), " ", chars)))
    ) %>%
    transmute(sample, chars, method, metric, value = p_num)
}

plot_data_direct_plus_method_eval<-function(data_long){
sum_dat <- data_long %>%
  group_by(chars, method, metric) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            sd   = sd(value,   na.rm = TRUE),
            n    = sum(!is.na(value)),
            .groups = "drop") %>%
  mutate(fill_group = paste(method, metric))

dat_norm <- data_long %>%
  mutate(
    metric = case_when(
      str_detect(metric, regex("^pg",  TRUE)) ~ "PGs",
      str_detect(metric, regex("^pre", TRUE)) ~ "Precursors",
      TRUE ~ metric
    ),
    method = factor(method, levels = c("Method evaluation","DirectDIA+")),
    chars  = factor(chars)
  )

pgs_tbl <- sum_dat %>% filter(metric=="PGs") %>%
  mutate(y = mean, lab_pgs = sprintf("%d", round(mean)))
pgs_pts <- filter(dat_norm, metric == "PGs")
prec_tbl <- sum_dat %>% filter(metric=="Precursors") %>%
  mutate(y = mean, lab_prec = sprintf("%.1f", mean))
prec_pts <- filter(dat_norm, metric == "Precursors")
p1<-ggplot() + 
  geom_col(data = pgs_tbl, aes(x = chars, y = y, group = method, fill = method),
           position = "identity",
           alpha = 0.7, 
           #position = pd,
           width = 0.72, 
           color = "black",
           linewidth = 0.25)+
    geom_line(data = pgs_tbl,aes(x=chars, y=y, group=method,color=method), position="identity", size=0.8) +
  scale_fill_manual(values = c("Method evaluation" = group_valuess[1], "DirectDIA+" = "#B0E0E6"))+
  scale_color_manual(values = c("Method evaluation" = group_valuess[1], "DirectDIA+" = "#B0E0E6"))+
  geom_text(data = pgs_tbl, aes(x = chars, y = y, label = scales::comma(round(y)), group = method),
            position = "identity",color = "black",
            fontface = "bold", size = 3.1, vjust = -0.5)+
  geom_point(data = pgs_pts, aes(x = chars, y = value, group = method),
             position =  "identity", shape = 21, size = 2, stroke = 0.2,
             fill = "white", color = "black", alpha = 0.35) +
  labs(x=NULL,y="Protein Groups")+
  theme_yzc

p2<-ggplot() + 
  geom_col(data = prec_tbl, aes(x = chars, y = y, group = method, fill = method),
           position = "identity",
           alpha = 0.7, 
           #position = pd,
           width = 0.72, 
           color = "black",
           linewidth = 0.25)+
    geom_line(data = prec_tbl,aes(x=chars, y=y, group=method,color=method), position="identity", size=0.8) +
  scale_fill_manual(values = c("Method evaluation" = group_valuess[2], "DirectDIA+" = "#DDA0DD"))+
    scale_color_manual(values = c("Method evaluation" = group_valuess[2], "DirectDIA+" = "#DDA0DD"))+
  geom_text(data = prec_tbl, aes(x = chars, y = y, label = scales::comma(round(y)), group = method),
            position = "identity",color = "black",
            fontface = "bold", size = 3.1, vjust = -0.2)+
  geom_point(data = prec_pts, aes(x = chars, y = value, group = method),
             position =  "identity", shape = 21, size = 2, stroke = 0.2,
             fill = "white", color = "black", alpha = 0.35) +
   #geom_line(data = prec_tbl,aes(x=chars, y=y, group=method,color=method), position="identity", size=0.8) +
  labs(x=NULL,y="Peptides")+
  theme_yzc
     
return(p1/p2)

}




spectronaut_amount_to_long <- function(data, feature_col = NULL) {
 #   data=data_pep_directdia_amoun
 # feature_col = "PEP.GroupingKey"
  qty <- data[,c(1,grep("\\.Quantity$", colnames(data)))]
  colnames(qty) <- c(feature_col,unname(sapply(colnames(qty[-1]), function(x) strsplit(x, "\\.")[[1]][4])))
  qty[qty == "NaN"] <- NA
  qty <- as.data.frame(qty)
  
  # Determine feature id for rows
  if (!is.null(feature_col) && feature_col %in% names(data)) {
    feature <- data[[feature_col]]
  } else if ("PG.ProteinGroups" %in% names(data)) {
    feature <- data[["PG.ProteinGroups"]]
  } else if ("EG.PrecursorId" %in% names(data)) {
    feature <- data[["EG.PrecursorId"]]
  } else if ("Precursor.Id" %in% names(data)) {
    feature <- data[["Precursor.Id"]]
  } else {
    feature <- paste0("feat_", seq_len(nrow(qty)))
  }
  
   a<-qty %>%
    tibble::rownames_to_column(".row") %>%
     #mutate(feature=feature)%>%
    select(-.row) %>%
    pivot_longer(-!!sym(feature_col), names_to = "sample", values_to = "intensity") %>%
     # as.data.frame(.)%>%
     dplyr::rename(feature =!!sym(feature_col)) %>%
    mutate(intensity = suppressWarnings(as.numeric(intensity)))
}
add_meta_amount <- function(df_long, method, metric, chars = "pg") {
  #method<-"DirectDIA+";metric="Precursors"
  pattern <- paste0("(?i)(\\d+)(?=\\s*", chars, ")")
  df_long %>%
    mutate(
      method = factor(method, levels = c("Method evaluation", "DirectDIA+")),
      metric = metric,
      chars_num = coalesce(
        suppressWarnings(as.integer(str_extract(sample, pattern))),
        suppressWarnings(as.integer(str_extract(sample, "\\\
\\d+")))
      ),
      chars = factor(paste0(chars_num, " ", chars),
                     levels = paste0(sort(unique(chars_num)), " ", chars))
    )
}
cv_from_amount <- function(df_long_meta, na_threshold = 0.5, min_reps = 2) {
  # total replicates per group (for NA filtering)
  #df_long_meta<-cv_long_cv
  n_total <- df_long_meta %>%
    group_by(chars, method, metric) %>%
    summarise(total = dplyr::n_distinct(sample), .groups = "drop")
  
  df_long_meta %>%
    group_by(chars, method, metric, feature) %>%
    summarise(
      n_non_na = sum(!is.na(intensity)),
      mean_val = mean(intensity, na.rm = TRUE),
      sd_val   = sd(intensity, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    left_join(n_total, by = c("chars","method","metric")) %>%
    mutate(
      prop_non_na = n_non_na / pmax(total, 1),
      cv = sd_val / mean_val
    ) %>%
    filter(total >= min_reps, n_non_na >= min_reps, prop_non_na >= (1 - na_threshold), is.finite(cv))
}
plot_cv_half_violin<-function(cv_tbl){
    cv_tbl <- cv_tbl %>% mutate(method = factor(method, levels = c("Method evaluation","DirectDIA+")))
  # Method evaluation only
    #cv_tbl<-cv_tbl_cv
cv_tbl_Pgs<-dplyr::filter(cv_tbl, metric == "PGs")
cv_tbl_Prec<-dplyr::filter(cv_tbl, metric != "PGs")
med_tbl_Pgs <- cv_tbl_Pgs %>%
  group_by(chars, method) %>%
  summarise(med = median(cv, na.rm = TRUE), .groups = "drop")

med_tbl_Prec <- cv_tbl_Prec %>%
  group_by(chars, method) %>%
  summarise(med = median(cv, na.rm = TRUE), .groups = "drop")

p1<-ggplot(cv_tbl_Pgs) +
  gghalves::geom_half_violin(
    aes(x = chars, y = cv*100, fill = method),
    data = cv_tbl_Pgs %>% filter(method == "DirectDIA+"),
    side = "l", color = "black", width = 0.9, trim = TRUE, alpha = 0.7
  ) +
  gghalves::geom_half_violin(
    aes(x = chars, y = cv*100, fill = method),
    data = cv_tbl_Pgs %>% filter(method != "DirectDIA+"),
    side = "r", color = "black", width = 0.9, trim = TRUE, alpha = 0.7
  ) +
  geom_text(
    data = med_tbl_Pgs,
    aes(x = chars, y = med*100, label = sprintf("%.2f", med*100), color = method),position = position_dodge(width = 0.45),
    vjust = 1, fontface = "bold", size = 3.5, show.legend = FALSE
  ) +
  ylim(0,50)+
  scale_fill_manual(values = c("DirectDIA+" = group_valuess[1], "Method evaluation" = group_valuess[2])) +
  scale_color_manual(values = c("DirectDIA+" = "white", "Method evaluation" = "white")) +
  labs(x = NULL, y = "Protein Groups CV (%)") +
 theme_yzc
p2<-ggplot(cv_tbl_Prec) +
  gghalves::geom_half_violin(
    aes(x = chars, y = cv*100, fill = method),
    data = cv_tbl_Prec %>% filter(method == "DirectDIA+"),
    side = "l", color = "black", width = 0.9, trim = TRUE, alpha = 0.7
  ) +
  gghalves::geom_half_violin(
    aes(x = chars, y = cv*100, fill = method),
    data = cv_tbl_Prec %>% filter(method != "DirectDIA+"),
    side = "r", color = "black", width = 0.9, trim = TRUE, alpha = 0.7
  ) +
  geom_text(
    data = med_tbl_Prec,
    aes(x = chars, y = med*100, label = sprintf("%.2f", med*100), color = method),position = position_dodge(width = 0.45),
    vjust = 1, fontface = "bold", size = 3.5, show.legend = FALSE
  ) +
  ylim(0,50)+
  scale_fill_manual(values = c("DirectDIA+" = group_valuess[1], "Method evaluation" = group_valuess[2])) +
  scale_color_manual(values = c("DirectDIA+" = "white", "Method evaluation" = "white")) +
  labs(x = NULL, y = "Peptides CV (%)") +
  theme_yzc

return(p1/p2)

  
}


```

```{r amount analysis}
data_pep_directdia_amoun<-read.delim("./Figure2_data/amount_directDIA.tsv")
data_pro_directdia_amoun<-read.delim("./Figure2_data/amount_PGs_directDIA.tsv")
data_pro_method_eval_amoun<-read.delim("./Figure2_data/amount_PGs_method_evaluation.tsv")
data_pep_method_eval_amoun<-read.delim("./Figure2_data/amount_method_evaluation.tsv")
  pep_directdia_amoun<-spectronaut_to_iden(data_pep_directdia_amoun)
  pro_directdia_amoun<-spectronaut_to_iden(data_pro_directdia_amoun)
  pep_method_eval_amoun<-spectronaut_to_iden(data_pep_method_eval_amoun)
  pro_method_eval_amoun<-spectronaut_to_iden(data_pro_method_eval_amoun)
dat_long_amoun <- bind_rows(
  add_meta(pep_directdia_amoun,   "DirectDIA+",        "Precursors", chars="pg"),
  add_meta(pro_directdia_amoun,   "DirectDIA+",        "PGs",        chars="pg"),
  add_meta(pep_method_eval_amoun, "Method evaluation", "Precursors", chars="pg"),
  add_meta(pro_method_eval_amoun, "Method evaluation", "PGs",        chars="pg")
)
p<-plot_data_direct_plus_method_eval(dat_long_amoun)
pdf("./Figure2_plot/Figure2C_amount.pdf",width=7,height=5)
(p[[1]]+ylim(0,6500))/(p[[2]]+ylim(0,52000))
dev.off()
cv_long_amoun <- bind_rows(
  spectronaut_amount_to_long(data_pep_directdia_amoun,feature_col = "PEP.GroupingKey")   %>% add_meta_amount("DirectDIA+",        "Precursors", chars = "pg"),
  spectronaut_amount_to_long(data_pro_directdia_amoun,feature_col = "PG.ProteinGroups")   %>% add_meta_amount("DirectDIA+",        "PGs",        chars = "pg"),
  spectronaut_amount_to_long(data_pep_method_eval_amoun,feature_col = "PEP.GroupingKey") %>% add_meta_amount("Method evaluation", "Precursors", chars = "pg"),
  spectronaut_amount_to_long(data_pro_method_eval_amoun,feature_col = "PG.ProteinGroups") %>% add_meta_amount("Method evaluation", "PGs",        chars = "pg")
)
cv_tbl_amount <- cv_from_amount(cv_long_amoun, na_threshold = 0.5, min_reps = 2)
p_cv_amount <- plot_cv_half_violin(cv_tbl_amount)
pdf("./Figure2_plot/Figure2C_amount_cv.pdf",width=7,height=5)
p_cv_amount
dev.off()


sum_dat_amoun <- dat_long_amoun %>%
  group_by(chars, method, metric) %>%
  summarise(mean = mean(value, na.rm = TRUE),
            sd   = sd(value,   na.rm = TRUE),
            n    = sum(!is.na(value)),
            .groups = "drop") %>%
  mutate(fill_group = paste(method, metric))

sum_dat_pep_amoun_DirectDIA <- sum_dat_amoun %>% filter(metric == "Precursors") %>% filter(method == "DirectDIA+") %>%
  arrange(desc(mean)) %>%
  mutate(chars = factor(chars, levels = chars)) %>%
  mutate(amount=as.numeric(strsplit(as.character(chars),"pg")[[1]][1]))


```


```{r CV analysis}
data_pep_directdia_cv<-read.delim("./Figure2_data/CV_pep_directdia.tsv")
data_pro_directdia_cv<-read.delim("./Figure2_data/CV_pro_directdia.tsv")
data_pro_method_eval_cv<-read.delim("./Figure2_data/CV_pro_method_eva.tsv")
data_pep_method_eval_cv<-read.delim("./Figure2_data/CV_pep_method_eva.tsv")
  pep_directdia_cv<-spectronaut_to_iden(data_pep_directdia_cv)
  pro_directdia_cv<-spectronaut_to_iden(data_pro_directdia_cv)
  pep_method_eval_cv<-spectronaut_to_iden(data_pep_method_eval_cv)
  pro_method_eval_cv<-spectronaut_to_iden(data_pro_method_eval_cv)
dat_long_cv <- bind_rows(
  add_meta(pep_directdia_cv,   "DirectDIA+",        "Precursors", chars="neg"),
  add_meta(pro_directdia_cv,   "DirectDIA+",        "PGs",        chars="neg"),
  add_meta(pep_method_eval_cv, "Method evaluation", "Precursors", chars="neg"),
  add_meta(pro_method_eval_cv, "Method evaluation", "PGs",        chars="neg")
)

p_cv<-plot_data_direct_plus_method_eval(dat_long_cv)
pdf("./Figure2_plot/Figure2C_cv.pdf",width=7,height=5)
(p_cv[[1]]+ylim(0,5000))/(p_cv[[2]]+ylim(0,36000))
dev.off()
cv_long_cv <- bind_rows(
  spectronaut_amount_to_long(data_pep_directdia_cv,feature_col = "PEP.GroupingKey")   %>% add_meta_amount("DirectDIA+",        "Precursors", chars = "neg"),
  spectronaut_amount_to_long(data_pro_directdia_cv,feature_col = "PG.ProteinGroups")   %>% add_meta_amount("DirectDIA+",        "PGs",        chars = "neg"),
  spectronaut_amount_to_long(data_pep_method_eval_cv,feature_col = "PEP.GroupingKey") %>% add_meta_amount("Method evaluation", "Precursors", chars = "neg"),
  spectronaut_amount_to_long(data_pro_method_eval_cv,feature_col = "PG.ProteinGroups") %>% add_meta_amount("Method evaluation", "PGs",        chars = "neg")
)
cv_tbl_cv <- cv_from_amount(cv_long_cv, na_threshold = 0.5, min_reps = 2)
p_cv_cv <- plot_cv_half_violin(cv_tbl_cv)
pdf("./Figure2_plot/Figure2C_cv_cv.pdf",width=7,height=5)
p_cv_cv
dev.off()
```


```{r Th analysis}
data_pep_directdia_Th<-read.delim("./Figure2_data/TH_directDIA.tsv")
data_pro_directdia_Th<-read.delim("./Figure2_data/TH_PGs_directDIA.tsv")
data_pro_method_eval_Th<-read.delim("./Figure2_data/TH_PGs_method_evaluation.tsv")
data_pep_method_eval_Th<-read.delim("./Figure2_data/TH_method_evaluation.tsv")
  pep_directdia_Th<-spectronaut_to_iden(data_pep_directdia_Th)
  pro_directdia_Th<-spectronaut_to_iden(data_pro_directdia_Th)
  pep_method_eval_Th<-spectronaut_to_iden(data_pep_method_eval_Th)
  pro_method_eval_Th<-spectronaut_to_iden(data_pro_method_eval_Th)
dat_long_Th <- bind_rows(
  add_meta(pep_directdia_Th,   "DirectDIA+",        "Precursors", chars="Th"),
  add_meta(pro_directdia_Th,   "DirectDIA+",        "PGs",        chars="Th"),
  add_meta(pep_method_eval_Th, "Method evaluation", "Precursors", chars="Th"),
  add_meta(pro_method_eval_Th, "Method evaluation", "PGs",        chars="Th")
)
p_Th<-plot_data_direct_plus_method_eval(dat_long_Th)
pdf("./Figure2_plot/Figure2C_Th.pdf",width=7,height=5)
(p_Th[[1]]+ylim(0,5000))/(p_Th[[2]]+ylim(0,32000))
dev.off()


cv_long_Th <- bind_rows(
  spectronaut_amount_to_long(data_pep_directdia_Th,feature_col = "PEP.GroupingKey")   %>% add_meta_amount("DirectDIA+",        "Precursors", chars = "Th"),
  spectronaut_amount_to_long(data_pro_directdia_Th,feature_col = "PG.ProteinGroups")   %>% add_meta_amount("DirectDIA+",        "PGs",        chars = "Th"),
  spectronaut_amount_to_long(data_pep_method_eval_Th,feature_col = "PEP.GroupingKey") %>% add_meta_amount("Method evaluation", "Precursors", chars = "Th"),
  spectronaut_amount_to_long(data_pro_method_eval_Th,feature_col = "PG.ProteinGroups") %>% add_meta_amount("Method evaluation", "PGs",        chars = "Th")
)
cv_tbl_Th <- cv_from_amount(cv_long_Th, na_threshold = 0.5, min_reps = 2)
p_cv_Th <- plot_cv_half_violin(cv_tbl_Th)
pdf("./Figure2_plot/Figure2C_Th_cv.pdf",width=7,height=5)
p_cv_Th
dev.off()

```




```{r one vs two}
data_one_vs_two_pro<-read.delim("./Figure3_data/20250919_104403_20250907_Report_pro.tsv")

# 
# data_one_vs_two_pro[5,2]<-"ARHGEF5"
# data_one_vs_two_pro[57,2]<-"KRT-85"
# data_one_vs_two_pro[42,2]<-"Tryspin"
# data_one_vs_two_pro[54,2]<-"cRAP-P15252"
# data_one_vs_two_pro[56,2]<-"cRAP-P22629"
# data_one_vs_two_pro[2606,2]<-"Q6ZSR9"

data_one_vs_two_pep<-read.delim("./Figure3_data/20250919_104403_20250907_Report_pep.tsv")

# pro_one_vs_two<-spectronaut_to_iden(data_one_vs_two_pro)
# 
# pro_one_vs_two$source<-c(rep("One-step",10),rep("Two-step",10))
# 
# pep_one_vs_two<-spectronaut_to_iden(data_one_vs_two_pep)
# pep_one_vs_two$source<-c(rep("One-step",10),rep("Two-step",10))

pro_one_vs_two<-spectronaut_to_iden(data_one_vs_two_pro)
pro_one_vs_two1<-data.frame(p_num=c(pro_one_vs_two$p_num,1685,2678,2376,1946,1901,2033,2719,1626,2171,2343,2029,2434,2708,2347,2729,1882,2089,2935))
pro_one_vs_two1$source<-c(rep("One-step",10),rep("Two-step",10),rep("yzc-2024",18))

pep_one_vs_two<-spectronaut_to_iden(data_one_vs_two_pep)
pep_one_vs_two1<-data.frame(p_num=c(pep_one_vs_two$p_num,5388,9148,7743,6109,5910,6594,9281,4476,7411,8135,6542,7967,9202,7864,9632,6106,7170,10740))
pep_one_vs_two1$source<-c(rep("One-step",10),rep("Two-step",10),rep("yzc-2024",18))




sum_pro_one_vs_two <- pro_one_vs_two1 %>%
  group_by(source) %>%
  summarise(mean = mean(p_num, na.rm = TRUE),
            sd   = sd(p_num,   na.rm = TRUE),
            n    = sum(!is.na(p_num)),
            .groups = "drop") %>%
  mutate(fill_group = paste(source))
sum_pep_one_vs_two <- pep_one_vs_two1 %>%
  group_by(source) %>%
  summarise(mean = mean(p_num, na.rm = TRUE),
            sd   = sd(p_num,   na.rm = TRUE),
            n    = sum(!is.na(p_num)),
            .groups = "drop") %>%
  mutate(fill_group = paste(source))
p_one_vs_two_left<-ggplot() +
  geom_bar(data = sum_pro_one_vs_two,aes(x=source,y=mean,fill = source),stat = "identity",alpha=0.8,width = 0.3) +
    geom_errorbar(data = sum_pro_one_vs_two,aes(x=source,ymin=(mean-sd), ymax=(mean+sd)), width=0.2, position=position_dodge(0.7)) +
  #geom_jitter(width = 0.12, size = 2, alpha = 0.7, show.legend = FALSE) +

  geom_signif(
    data = pro_one_vs_two1, aes(x = source, y = p_num),
    comparisons = list(
      c("One-step", "Two-step"),
      c("One-step", "yzc-2024"),
      c("Two-step", "yzc-2024")
    ),
    test = "wilcox.test",
    map_signif_level = function(p) {
      stars <- ifelse(p < 0.001, "***",
               ifelse(p < 0.01,  "**",
               ifelse(p < 0.05,  "*", "ns")))
      #sprintf("p = %.8f\n%s", p, stars)
      sprintf("%s", stars)
    },
      y_position = c(
        max(pro_one_vs_two1$p_num) + 300,
        max(pro_one_vs_two1$p_num) + 600,
        max(pro_one_vs_two1$p_num) + 0
    ),
     tip_length = 0.01,
     textsize = 5
  )+
  ylim(0,5000)+
 
  scale_fill_manual(values = c("One-step" =group_valuess[[1]], "Two-step" = group_valuess[[2]],"yzc-2024"=group_valuess[[3]])) +
  labs(x = NULL, y = "Protein Groups") +
 theme_yzc+
  theme(legend.position = "none")

p_one_vs_two_right<-ggplot() +
  geom_bar(data = sum_pep_one_vs_two,aes(x=source,y=mean,fill = source),stat = "identity",alpha=0.8,width = 0.3) +
    geom_errorbar(data = sum_pep_one_vs_two,aes(x=source,ymin=(mean-sd), ymax=(mean+sd)), width=0.2, position=position_dodge(0.7)) +
  #geom_jitter(width = 0.12, size = 2, alpha = 0.7, show.legend = FALSE) +

  geom_signif(
    data = pep_one_vs_two1, aes(x = source, y = p_num),
    comparisons = list(
      c("One-step", "Two-step"),
      c("One-step", "yzc-2024"),
      c("Two-step", "yzc-2024")
    ),
    test = "wilcox.test",
    map_signif_level = function(p) {
      stars <- ifelse(p < 0.001, "***",
               ifelse(p < 0.01,  "**",
               ifelse(p < 0.05,  "*", "ns")))
      #sprintf("p = %.8f\n%s", p, stars)
      sprintf("%s", stars)
    },
      y_position = c(
        max(pep_one_vs_two1$p_num) + 1200,
        max(pep_one_vs_two1$p_num) + 2400,
        max(pep_one_vs_two1$p_num) + 0
    ),
     tip_length = 0.01,
     textsize = 5
  )+
  ylim(0,22000)+
 
  scale_fill_manual(values = c("One-step" =group_valuess[[1]], "Two-step" = group_valuess[[2]],"yzc-2024"=group_valuess[[3]])) +
  labs(x = NULL, y = "Peptides") +
 theme_yzc+
  theme(legend.position = "none")
pdf("./Figure2_plot/Figure2D_one_vs_two_iden.pdf",width = 12,height = 4)
p_one_vs_two_left + p_one_vs_two_right+p_one_vs_two_mid
dev.off()

pdf("./Figure3_plot/Figure3A_one_vs_two_corr.pdf",width = 5,height = 5)
scatter_plot_with_r_and_p_3_from_data_and_info_sp_pro(data_one_vs_two_pro,c(rep("One-step",10),rep("Two-step",10)))+
    scale_color_distiller(palette= 3, direction=1)+
  xlab("One-step Log10 Intensity")+
  ylab("Two-step Log10 Intensity")+
  theme_yzc+
  theme(legend.position = "none")

dev.off()


```


```{r one vs two DEP}
data_one_vs_two_pro<-data_one_vs_two_pro%>%  filter(!grepl("cRAP", PG.ProteinGroups)) 
data_clean_one_vs_two_pro_dep <- data_one_vs_two_pro %>%
  select(PG.Genes, matches("\\.PG\\.Quantity")) %>%
 dplyr::filter(PG.Genes != "")
data_clean_one_vs_two_pro_dep[data_clean_one_vs_two_pro_dep=="NaN"] <- NA
data_clean_one_vs_two_pro_dep<-keep_first_Gene_sp(data_clean_one_vs_two_pro_dep)
#prot_df_clean<-keep_first_Gene(prot_df_clean)
# data_clean_one_vs_two_pro_dep<-samerow_mean(data_clean_one_vs_two_pro_dep)
rownames(data_clean_one_vs_two_pro_dep)<-data_clean_one_vs_two_pro_dep$PG.Genes
data_clean_one_vs_two_pro_dep<-data_clean_one_vs_two_pro_dep[,-1]

group_info_one_vs_two_pro_dep<-data.frame(group=c(rep("One-step",10),rep("Two-step",10)))
group_info_one_vs_two_pro_dep$sample<-colnames(data_clean_one_vs_two_pro_dep)
group_info_one_vs_two_pro_dep<-group_info_one_vs_two_pro_dep %>% dplyr::select(sample,group)
rownames(group_info_one_vs_two_pro_dep)<-group_info_one_vs_two_pro_dep$sample
data_clean_one_vs_two_pro_dep_filter<-filter_na_bygroup(data_clean_one_vs_two_pro_dep, group_info_one_vs_two_pro_dep , x = 0.5)
data_clean_one_vs_two_pro_dep_filter<-median_correct(data_clean_one_vs_two_pro_dep_filter)
data_clean_one_vs_two_pro_dep_filter1<-data_clean_one_vs_two_pro_dep_filter
library(impute)
data_clean_one_vs_two_pro_dep_filter1<-impute.knn(as.matrix(log2(data_clean_one_vs_two_pro_dep_filter)) ,k = 5, rowmax = 0.7, colmax = 0.8, maxp = 1500, rng.seed=362436069)
data_clean_one_vs_two_pro_dep_filter1<-data.frame(data_clean_one_vs_two_pro_dep_filter1$data)
colnames(data_clean_one_vs_two_pro_dep_filter1)<-colnames(data_clean_one_vs_two_pro_dep_filter)
DEP_one_vs_two_pro<-limma_nopair_test_2cluster(data_log2 = log2(data_clean_one_vs_two_pro_dep_filter),cluster = group_info_one_vs_two_pro_dep%>%dplyr::select(group),desc = F,p_value = 0.01,Fc_value =1.5,method = "BH")
vocanol_plot(DEP_one_vs_two_pro$data,p.value = 0.01,Fc_value=1.5)+theme_yzc







library(dplyr)
expr_mat<-data_clean_one_vs_two_pro_dep
meta<-group_info_one_vs_two_pro_dep
expr_mat <- expr_mat[, meta$sample]

expr_bin <- !is.na(expr_mat)

groups <- unique(meta$group)

feature_list <- lapply(groups, function(g) {
  
  samples_g <- meta %>%
    filter(group == g) %>%
    pull(sample)
  
  mat_g <- expr_bin[, samples_g]
  
  ratio <- rowMeans(mat_g, na.rm = TRUE)
  
  # ≥50%
  features <- names(ratio[ratio >= 0.5])
  
  return(features)
})

names(feature_list) <- groups
group1 <- feature_list[[1]]
group2 <- feature_list[[2]]

common_features <- intersect(group1, group2)
unique_g1 <- setdiff(group1, group2)
unique_g2 <- setdiff(group2, group1)
library(ggVennDiagram)

venn <- ggVennDiagram(feature_list, label_alpha = 0) +
  scale_fill_gradient(low = "#F4FAFE", high = "#4981BF") +
  theme_void()+coord_flip()+theme_yzc


pdf("./Figure2_plot/Figure2F_veen_plot.pdf",width = 5,height = 4)
print(venn)
dev.off()


library(dplyr)
library(ggplot2)
library(tibble)

group1_name <- "One-step"
group2_name <- "Two-step"
pseudo <- 1

meta <- meta %>% filter(sample %in% colnames(expr_mat))
expr_mat <- expr_mat[, meta$sample, drop = FALSE]

get_group_rank_df <- function(expr_mat, meta, group_name, detected_features = NULL, pseudo = 1) {
  samples_g <- meta %>%
    filter(group == group_name) %>%
    pull(sample)
  
  mat_g <- expr_mat[, samples_g, drop = FALSE]
  
  if (!is.null(detected_features)) {
    mat_g <- mat_g[detected_features, , drop = FALSE]
  }
  
  intensity <- apply(mat_g, 1, function(x) {
    if (all(is.na(x))) return(NA_real_)
    median(x, na.rm = TRUE)
  })
  
  df <- tibble(
    feature = rownames(mat_g),
    intensity = intensity
  ) %>%
    filter(!is.na(intensity), intensity > 0) %>%
    arrange(desc(intensity)) %>%
    mutate(
      rank = row_number(),
      log10_intensity = log10(intensity + pseudo),
      group = group_name
    )
  
  df
}

df_g1 <- get_group_rank_df(
  expr_mat, meta, group1_name,
  detected_features = feature_list[[group1_name]],
  pseudo = pseudo
)

df_g2 <- get_group_rank_df(
  expr_mat, meta, group2_name,
  detected_features = feature_list[[group2_name]],
  pseudo = pseudo
)

rank_df <- bind_rows(df_g1, df_g2) %>%
  mutate(
    highlight = ifelse(group == group1_name & feature %in% unique_g1,
                       "one-step unique", "other")
  )

p <- ggplot() +
  geom_line(
    data = rank_df,
    aes(x = rank, y = log10_intensity, color = group),
    linewidth = 0.9,
    alpha = 0.9
  ) +
  geom_point(
    data = rank_df %>% filter(highlight == "one-step unique"),
    aes(x = rank, y = log10_intensity),
    color = "red",
    size = 1.5,
    alpha = 0.6
  ) +
  labs(
    x = "Rank",
    y = expression(log[10](Intensity)),
    color = NULL
  ) +
  scale_color_manual(values = c("One-step" = group_valuess[[1]], "Two-step" = group_valuess[[2]])) +
  theme_classic(base_size = 14)+theme_yzc
pdf("./Figure2_plot/Figure2F_rank_plot.pdf",width = 5,height = 4)
print(p)
dev.off()

```

```{r protein_abs_copynum}
proteins_all <- read.csv("./ref_code/Protein_abs_copynum/HeLaHF14k.csv")
  rownames(proteins_all)<-proteins_all$Accession
  gene<-data.frame(proteins_all$Accession)
  print(dim(gene)[1])
  colnames(gene)<-"gene"
  
  
  
#all
data_pro_compare_all<- data_one_vs_two_pro %>%
  select(PG.ProteinGroups, matches("\\.PG\\.Quantity")) %>%
  #filter(grepl("cRAP", PG.ProteinGroups)) %>%
  pivot_longer(
    cols = matches("\\.PG\\.Quantity"),
    names_to = "Sample",
    values_to = "Quantity"
  )

data_pro_compare_all$Log10Quantity<-log10(data_pro_compare_all$Quantity+1)
group_info_compare_cRAP<-data.frame(Sample=unique(data_pro_compare_all$Sample),
                                 source=c(rep("One-step",10),rep("Two-step",10)))
data_pro_compare_all<-merge(data_pro_compare_all,group_info_compare_cRAP,by="Sample")

  sum_dat_compare <- data_pro_compare_all %>%
  group_by(source,PG.ProteinGroups) %>%
  summarise(mean = mean(Log10Quantity, na.rm = TRUE),
            sd   = sd(Log10Quantity,   na.rm = TRUE),
            n    = sum(!is.na(Log10Quantity)),
            .groups = "drop") %>%
  mutate(fill_group = source)
sum_dat_compare<-sum_dat_compare%>% filter(n>2)
plate_all<-unique(sum_dat_compare$PG.ProteinGroups[sum_dat_compare$source=="One-step"])
AM_DMF_SCP_all<-unique(sum_dat_compare$PG.ProteinGroups[sum_dat_compare$source=="Two-step"])

  ### Get the overlap
  proteins_both <- proteins_all %>%
    dplyr::filter( Accession%in% plate_all)
  proteins_both2 <- proteins_all %>%
    filter(Accession %in% AM_DMF_SCP_all)
  
  three<-listn[[1]]
  threeacc<-clusterProfiler::bitr(three, fromType = "SYMBOL", toType = "UNIPROT",OrgDb ="org.Hs.eg.db")
  proteins_both3<-proteins_all %>%
    filter(Accession %in% threeacc$UNIPROT)
proteins_all_final<-proteins_all
color_all<-"black"
color_both<-"#AF46B4"
transparency<-0.1
p_one_vs_two_mid<-ggplot(proteins_all_final) +
    xlab("Log")+
    ylab("Count")+
    geom_histogram(
      data = proteins_all_final,
      aes_string(x = "Log"), 
      color = color_all,
      fill = color_all,
      alpha = transparency,
      bins = 20
    ) +geom_vline(xintercept=mean(proteins_all_final$Log),linetype = "dashed")+
    geom_histogram(
      data = proteins_both,
      aes_string(x = "Log"), 
      color = "#4682B4",
      fill = "#4682B4",
      alpha = transparency,
      bins = 20
    )+geom_vline(xintercept=mean(proteins_both$Log),color="#4682B4",linetype = "dashed")+
    geom_histogram(
      data = proteins_both2,
      aes_string(x = "Log"), 
      color = "#AF46B4",
      fill = color_both,
      alpha = transparency,
      bins = 20
    ) +geom_vline(xintercept=mean(proteins_both2$Log),color="#AF46B4",linetype = "dashed")+
        geom_histogram(
      data = proteins_both3,
      aes_string(x = "Log"), 
      color = group_valuess[[3]],
      fill = color_both,
      alpha = transparency,
      bins = 20
    ) +geom_vline(xintercept=mean(proteins_both3$Log),color=group_valuess[[3]],linetype = "dashed")+
    
    geom_text(aes(x =4, y = 1000, label = paste0("Mean(All) = ", round(mean(proteins_all_final$Log), 2))),
                    hjust = 0, vjust = 1, color = "black", size = 6) +  
    geom_text(aes(x =4, y = 1000, label = paste0("\nMean(One-step)= ",round(mean(proteins_both$Log), 2))),
                    hjust = 0, vjust = 1, color = "#4682B4", size =6) +  
    geom_text(aes(x =4, y = 1000, label = paste0("\n\nMean(Two-step)= ",round(mean(proteins_both2$Log), 2))),
                    hjust = 0, vjust = 1, color = "#AF46B4", size = 6) +
      geom_text(aes(x =4, y = 1000, label = paste0("\n\n\nMean(yzc-2024)= ",round(mean(proteins_both3$Log), 2))),
                    hjust = 0, vjust = 1, color =   group_valuess[[3]], size = 6) +

  
    theme_bw() +
    
    theme(  #legend.position = "none",
            #axis.line = element_line(size=1.5),
            axis.text.x = element_text(color="black", size=16, face="plain",angle=0,hjust=0.5) ,# Values for face are one of "plain", "italic", "bold" and "bold.italic"
            axis.text.y = element_text(color="black", size=16, face="plain"),
            axis.title.x = element_text(color="black", size=16, face="bold"),
            axis.title.y = element_text(color="black", size=16, face="bold"),
            legend.text=element_text(color="black", size=16, face="bold"),
            legend.title=element_text(color="black", size=16, face="bold"),
            title = element_text(color="black", size=16, face="bold"))+
  theme_yzc

    
dev.off()

```

```{r corrmat}
library(PerformanceAnalytics)
data_one_vs_two_pro_1<- data_one_vs_two_pro %>%
  select(matches("\\.PG\\.Quantity"))
 colnames(data_one_vs_two_pro_1)<-c(paste("One",1:10),paste("Two",1:10))
panel.cor <- function(x, y, digits=2, prefix="", cex.cor) 
   {
       usr <- par("usr"); on.exit(par(usr)) 
       par(usr = c(0, 1, 0, 1))
      r <- abs(cor(x, y,use="pairwise.complete.obs")) 
       txt <- format(c(r, 0.123456789), digits=digits)[1] 
       txt <- paste(prefix, txt, sep="") 
       if(missing(cex.cor)) cex <- 0.8/strwidth(txt) 
       
           test <- cor.test(x,y) 
           # borrowed from printCoefmat
             Signif <- symnum(test$p.value, corr = FALSE, na = FALSE, 
                                                   cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
                                                  symbols = c("***", "**", "*", ".", " ")) 
             
               text(0.5, 0.5, txt, cex = cex * r) 
             text(.8, .8, Signif, cex=cex, col=2) 
           }
png("./Figure2_plot/Figure2E_corr.png",width = 7,height = 7)
 pairs(log2(data_one_vs_two_pro_1), upper.panel=panel.cor, gap=0,font.labels = 0.6)
 dev.off()

 
 df_med <- log10(data_one_vs_two_pro_1) |>
  rowwise() |>
  mutate(
    one_med = median(c_across(starts_with("One")), na.rm = TRUE),
    two_med = median(c_across(starts_with("Two")), na.rm = TRUE)
  ) |>
  ungroup() |>
  filter(!is.na(one_med), !is.na(two_med))

 library(ggplot2)
library(hexbin)

r_p <- cor(df_med$one_med, df_med$two_med, method = "pearson")
r_s <- cor(df_med$one_med, df_med$two_med, method = "spearman")

p_cor_2<-ggplot(df_med, aes(one_med, two_med)) +
  geom_hex(bins = 120) +
  scale_fill_viridis_c(trans = "log10") +
  geom_abline(slope = 1, intercept = 0,
              linetype = "dashed", color = "red") +
  annotate(
    "text",
    x = min(df_med$one_med) + 0.5,
    y = max(df_med$two_med) - 0.5,
    label = paste0(
      "Pearson r = ", round(r_p, 2), "\n"
     # "Spearman ρ = ", round(r_s, 2)
    ),
    size = 4.5
  ) +
  labs(
    x = "One-step log10 intensity",
    y = "Two-step log10 intensity"
  ) +
  theme_yzc
pdf("./Figure2_plot/Figure2E_corr_med.pdf",width = 7,height = 7)
p_cor_2+theme(legend.position = "none")
dev.off()

# 
#  library(pheatmap)
# 
# one_mat <-  log10(data_one_vs_two_pro_1) |>
#   select(starts_with("One")) |>
#   as.matrix()
# 
# cor_one <- cor(one_mat, use = "pairwise.complete.obs", method = "pearson")
# 
# pheatmap(
#   cor_one,
#   cluster_rows = FALSE,
#   cluster_cols = FALSE,
#   main = "One-step replicate correlation (Pearson)"
# )
df_ba <- df_med |>
  mutate(
    mean_logI = (one_med + two_med) / 2,
    diff_logI = one_med - two_med
  )
bias <- mean(df_ba$diff_logI, na.rm = TRUE)
sd_diff <- sd(df_ba$diff_logI, na.rm = TRUE)

loa_upper <- bias + 1.96 * sd_diff
loa_lower <- bias - 1.96 * sd_diff
library(ggplot2)

p_ba<-ggplot(df_ba, aes(mean_logI, diff_logI)) +
  geom_point(alpha = 0.35, size = 0.8) +
  geom_hline(yintercept = bias, color = "red", linewidth = 1) +
  geom_hline(yintercept = loa_upper, linetype = "dashed") +
  geom_hline(yintercept = loa_lower, linetype = "dashed") +
  annotate(
    "text",
    x = min(df_ba$mean_logI, na.rm = TRUE),
    y = bias,
    label = paste0("Bias = ", round(bias, 3)),
    hjust = 0,
    vjust = -0.5
  ) +
  annotate(
    "text",
    x = min(df_ba$mean_logI, na.rm = TRUE),
    y = loa_upper,
    label = paste0("+1.96 SD = ", round(loa_upper, 3)),
    hjust = 0,
    vjust = -0.5
  ) +
  annotate(
    "text",
    x = min(df_ba$mean_logI, na.rm = TRUE),
    y = loa_lower,
    label = paste0("-1.96 SD = ", round(loa_lower, 3)),
    hjust = 0,
    vjust = 1.2
  ) +
  labs(
    x = "Mean log10 intensity",
    y = "Difference (one-step − two-step)",
   # title = "Bland–Altman analysis of quantitative agreement"
  ) +
  theme_yzc

 
 
pdf("./Figure2_plot/Figure2E_bland_altman.pdf",width = 7,height = 7)
p_ba
dev.off()







df_mv <- log10(data_one_vs_two_pro_1) |>
  rowwise() |>
  mutate(
    one_med = median(c_across(starts_with("One")), na.rm = TRUE),
    two_med = median(c_across(starts_with("Two")), na.rm = TRUE),

    one_miss = mean(is.na(c_across(starts_with("One")))),
    two_miss = mean(is.na(c_across(starts_with("Two"))))
  ) |>
  ungroup()
df_bin <- df_mv |>
  filter(!is.na(one_med)) |>
  mutate(
    intensity_bin = cut(
      one_med,
      breaks = quantile(one_med, probs = seq(0, 1, by = 0.1), na.rm = TRUE),
      include.lowest = TRUE
    )
  ) |>
  group_by(intensity_bin) |>
  summarise(
    one_miss_rate = mean(one_miss, na.rm = TRUE),
    two_miss_rate = mean(two_miss, na.rm = TRUE),
    n_protein = n(),
    bin_center = mean(one_med, na.rm = TRUE)
  )

library(ggplot2)

ggplot(df_bin, aes(bin_center)) +
  geom_line(aes(y = one_miss_rate, color = "One-step"), linewidth = 1.2) +
  geom_point(aes(y = one_miss_rate, color = "One-step"), size = 2) +
  geom_line(aes(y = two_miss_rate, color = "Two-step"), linewidth = 1.2) +
  geom_point(aes(y = two_miss_rate, color = "Two-step"), size = 2) +
  scale_color_manual(values = c("One-step" = "#2C7FB8", "Two-step" = "#C77CFF")) +
  labs(
    x = "Protein abundance (log10 intensity)",
    y = "Missing value rate",
    color = "",
    title = "Missing value rate as a function of protein abundance"
  ) +
  theme_yzc



library(dplyr)
library(ggplot2)

df_consistency <- log10(data_one_vs_two_pro_1) |>
  rowwise() |>
  mutate(
    one_med = median(c_across(starts_with("One")), na.rm = TRUE),
    two_med = median(c_across(starts_with("Two")), na.rm = TRUE),

    one_detected = sum(!is.na(c_across(starts_with("One")))),
    two_detected = sum(!is.na(c_across(starts_with("Two"))))
  ) |>
  ungroup() |>
  filter(!is.na(one_med), !is.na(two_med))

df_bin <- df_consistency |>
  mutate(
    intensity_bin = cut(
      one_med,
      breaks = quantile(one_med, probs = seq(0, 1, 0.1), na.rm = TRUE),
      include.lowest = TRUE
    )
  ) |>
  group_by(intensity_bin) |>
summarise(
  bin_center = mean(one_med, na.rm = TRUE),
  one_detect_mean = mean(one_detected),
  one_detect_se = sd(one_detected) / sqrt(n()),
  two_detect_mean = mean(two_detected),
  two_detect_se = sd(two_detected) / sqrt(n())
)

# df_bin columns:
# bin_center, one_detect_mean, two_detect_mean

library(ggplot2)

p_rib<-ggplot(df_bin, aes(x = bin_center)) +
  geom_ribbon(
    aes(
      ymin = one_detect_mean - one_detect_se,
      ymax = one_detect_mean + one_detect_se
    ),
    fill = "#2C7FB8", alpha = 0.2
  ) +
  geom_line(
    aes(y = one_detect_mean, color = "One-step"),
    linewidth = 1.4
  ) +
  geom_point(
    aes(y = one_detect_mean, color = "One-step"),
    size = 2.8
  ) +
  geom_ribbon(
    aes(
      ymin = two_detect_mean - two_detect_se,
      ymax = two_detect_mean + two_detect_se
    ),
    fill = "#C77CFF", alpha = 0.2
  ) +
  geom_line(
    aes(y = two_detect_mean, color = "Two-step"),
    linewidth = 1.4
  ) +
  geom_point(
    aes(y = two_detect_mean, color = "Two-step"),
    size = 2.8
  ) +
  scale_color_manual(
    values = c("One-step" = "#2C7FB8", "Two-step" = "#C77CFF")
  ) +
  labs(
    x = "Protein abundance (log10 intensity)",
    y = "Detected replicates (out of 10)",
    color = ""
  ) +
  theme_yzc



pdf("./Figure2_plot/Figure2F_detection_consistency.pdf",width = 7,height = 7)
p_rib
dev.off()
```

```{r mis cleaveage}
data_one_vs_two_pep<-read.delim("./Figure3_data/20250919_104403_20250907_Report_pep.tsv")

data_one_vs_two_pep[data_one_vs_two_pep=="NaN"]<-NA
rownames(data_one_vs_two_pep)<-data_one_vs_two_pep$PEP.GroupingKey
data_one_vs_two_pep$PEP.GroupingKey<-NULL
df_miss_onevstwo<-calculate_miscleavage_per_sample(data_one_vs_two_pep) %>% filter(Miscleavage %in% 1)
df_miss_onevstwo$Group1<-unname(sapply(df_miss_onevstwo$Sample, function(x) strsplit(x, "_")[[1]][5]))
df_miss_onevstwo$Group1<-factor(df_miss_onevstwo$Group1, levels=c("onestep","twostep"))
# df_miss_onevstwo$Group2<-unname(sapply(df_miss_onevstwo$Sample, function(x) strsplit(x, "_")[[1]][5]))
# df_miss_onevstwo$Group2<-factor(df_miss_onevstwo$Group2, levels=c("5min","15min","30min","60min","120min"))
sum_dat <- df_miss_onevstwo %>%
  group_by(Group1) %>%
  summarise(mean = mean(Fraction, na.rm = TRUE),
            sd   = sd(Fraction,   na.rm = TRUE),
            n    = sum(!is.na(Fraction)),
            .groups = "drop") %>%
  mutate(fill_group = paste(Group1))
step_mis_compare<-ggplot(sum_dat, aes(x=Group1, y=mean*200,fill=Group1)) +
  geom_bar(stat="identity", position=position_dodge(), width=0.7,alpha=0.8) +
  geom_errorbar(aes(ymin=(mean-sd)*200, ymax=(mean+sd)*200), width=0.2, position=position_dodge(0.7),alpha=0.8) +
   geom_signif(
    data = df_miss_onevstwo, aes(x = Group1, y = Fraction),
    comparisons = list(
      c("onestep", "twostep")
    ),
    test = "t.test",
    map_signif_level = function(p) {
      stars <- ifelse(p < 0.001, "***",
               ifelse(p < 0.01,  "**",
               ifelse(p < 0.05,  "*", "ns")))
      #sprintf("p = %.8f\n%s", p, stars)
      sprintf("%s", stars)
    },
      y_position = c(
        max(df_miss_onevstwo$Fraction)*200 + 2
    
    ),
     tip_length = 0.01,
     textsize = 5
  )+

  geom_text(data = sum_dat, aes(x = Group1, y = mean*200, label = scales::comma(round(mean*200,4)), group = Group1),
            position = position_dodge(0.7),color = "white",
            fontface = "bold", size = 4, vjust = 5)+
  geom_point(data = df_miss_onevstwo, aes(x = Group1, y = Fraction*200, group = Group1),
             position =  position_dodge(0.7), shape = 21, size = 2, stroke = 0.2,
             fill = "white", color = "black", alpha = 0.85) +
  #scale_fill_manual(values=group_valuess,guide = "none") +
  scale_fill_manual(values=group_valuess) +
  labs(x="", y="Percentage of Missed Cleavages (%)", fill="Group")+
  theme_yzc
pdf("./Figure3_plot/Figure3C_miscleavage_compare.pdf",width = 5,height = 7)
step_mis_compare
dev.off()






 data_one_vs_two_pep_long<- spectronaut_amount_to_long(data_one_vs_two_pep,feature_col = "PEP.GroupingKey") 
data_one_vs_two_pro_long<-spectronaut_amount_to_long(data_one_vs_two_pro,feature_col = "PG.ProteinGroups")

 data_one_vs_two_pep_long<- data_one_vs_two_pep_long %>% mutate(group = unname(sapply(sample, function(x) strsplit(x, "_")[[1]][5])))
 data_one_vs_two_pro_long<- data_one_vs_two_pro_long %>% mutate(group = unname(sapply(sample, function(x) strsplit(x, "_")[[1]][5])))

na_threshold = 0.5
min_reps = 3
  n_total <- data_one_vs_two_pep_long %>%
    group_by(group) %>%
    summarise(total = dplyr::n_distinct(sample), .groups = "drop")
  
 data_one_vs_two_pep_long_cv<- data_one_vs_two_pep_long %>%
    group_by(feature,group) %>%
    summarise(
      n_non_na = sum(!is.na(intensity)),
      mean_val = mean(intensity, na.rm = TRUE),
      sd_val   = sd(intensity, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    left_join(n_total) %>%
    mutate(
      prop_non_na = n_non_na / pmax(total, 1),
      cv = sd_val / mean_val
    ) %>%
    filter(total >= min_reps, n_non_na >= min_reps, prop_non_na >= (1 - na_threshold), is.finite(cv))
 
   n_total <- data_one_vs_two_pro_long %>%
    group_by(group) %>%
    summarise(total = dplyr::n_distinct(sample), .groups = "drop")
  
 data_one_vs_two_pro_long_cv<- data_one_vs_two_pro_long %>%
    group_by(feature,group) %>%
    summarise(
      n_non_na = sum(!is.na(intensity)),
      mean_val = mean(intensity, na.rm = TRUE),
      sd_val   = sd(intensity, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    left_join(n_total) %>%
    mutate(
      prop_non_na = n_non_na / pmax(total, 1),
      cv = sd_val / mean_val
    ) %>%
    filter(total >= min_reps, n_non_na >= min_reps, prop_non_na >= (1 - na_threshold), is.finite(cv))
 
 
 
 data_one_vs_two_pro_long_cv$metric<-"PGs"
 data_one_vs_two_pep_long_cv$metric<-"PEPs"
 
  cv_tbl <- rbind(data_one_vs_two_pro_long_cv, data_one_vs_two_pep_long_cv)
  # Method evaluation only
    #cv_tbl<-cv_tbl_cv
cv_tbl_Pgs<-dplyr::filter(cv_tbl, metric == "PGs")
cv_tbl_Prec<-dplyr::filter(cv_tbl, metric != "PGs")
med_tbl_Pgs <- cv_tbl_Pgs %>%
  group_by(group) %>%
  summarise(med = median(cv, na.rm = TRUE), .groups = "drop")
med_tbl_Pgs$metric<-"PGs"
med_tbl_Prec <- cv_tbl_Prec %>%
  group_by(group) %>%
  summarise(med = median(cv, na.rm = TRUE), .groups = "drop")
med_tbl_Prec$metric<-"PEPs"
med_tbl <- rbind(med_tbl_Pgs, med_tbl_Prec)
pdf("./Figure2_plot/Figure2D_cv_compare1.pdf",width = 7,height = 7)
ggplot(cv_tbl) +
  gghalves::geom_half_violin(
    aes(x = group, y = cv*100, fill = metric),
    data = cv_tbl %>% filter(metric == "PGs"),
    side = "l", color = "black", width = 0.9, trim = TRUE, alpha = 0.7
  ) +
  gghalves::geom_half_violin(
    aes(x = group, y = cv*100, fill = metric),
    data = cv_tbl %>% filter(metric != "PGs"),
    side = "r", color = "black", width = 0.9, trim = TRUE, alpha = 0.7
  ) +
  geom_text(
    data = med_tbl,
    aes(x = group, y = med*100, label = sprintf("%.2f", med*100), color = "black"),position = position_dodge(width = 0.45),
    vjust = 1, fontface = "bold", size = 3.5, show.legend = FALSE
  ) +
  ylim(0,100)+
  scale_fill_manual(values = c("PEPs" = group_valuess[1], "PGs" = group_valuess[2])) +
  labs(x = NULL, y = "CV (%)") +
 theme_yzc
dev.off()


















fasta_file_path<-"/Users/zhichengyang/Downloads/UP000005640_9606_Homo_sapiens_20210617_con.fasta_td.fasta"

get_hydrovalue_kyte_doolittle_genelist<-function(fasta_file_path,geneList){
get_seq_from_fasta_genelist<-function(fasta_file_path=fasta_file_path,geneList=c("TP53")){
  library(Biostrings)
  library(dplyr)
  sequences <- readBStringSet(fasta_file_path)
  sq<-data.frame(sequences)
  sq$name<-rownames(sq)
  se1<-sq
  extract_gene_name <- function(str) {
    return(sub(".*GN=([^ ]+).*", "\\1", str))
  }
  
  gene_names <- sapply(sq$name, extract_gene_name)
  se1$gene_names<-gene_names
  se2<-se1[se1$gene_names %in% geneList,]
  
  
  return(se2)
}
a<-get_seq_from_fasta_genelist(fasta_file_path=fasta_file_path,geneList =geneList)
hydrophobicity_kyte_doolittle <- list(
  'A' = 1.8,  'C' = 2.5,  'D' = -3.5, 'E' = -3.5, 
  'F' = 2.8,  'G' = -0.4, 'H' = -3.2, 'I' = 4.5,
  'K' = -3.9, 'L' = 3.8,  'M' = 1.9,  'N' = -3.5,
  'P' = -1.6, 'Q' = -3.5, 'R' = -4.5, 'S' = -0.8,
  'T' = -0.7, 'V' = 4.2,  'W' = -0.9, 'Y' = -1.3
)
calculate_protein_hydrophobicity <- function(protein_sequence) {
  total_score <- 0
  for (aa in strsplit(protein_sequence, '')[[1]]) {
    if (aa %in% names(hydrophobicity_kyte_doolittle)) {
      total_score <- total_score + hydrophobicity_kyte_doolittle[[aa]]
    }
  }
  return(total_score / nchar(protein_sequence))
}
results <- sapply(a$sequences, calculate_protein_hydrophobicity)
value<-data.frame(results)

return(value)
}





one<-get_hydrovalue_kyte_doolittle_genelist(fasta_file_path,feature_list[[1]])
two<-get_hydrovalue_kyte_doolittle_genelist(fasta_file_path,feature_list[[2]])




one_df <- data.frame(Value = one[,1], Group = "onestep")
two_df <- data.frame(Value = two[,1], Group = "twostep")

sum_dat <- bind_rows(one_df, two_df)

library(ggplot2)
library(ggsignif)


pdf("./Figure2_plot/Figure2D_hydrovalue_compare1.pdf",width = 7,height = 7)

ggplot(sum_dat, aes(x = Group, y = Value, fill = Group)) +
  
  geom_violin(trim = FALSE, alpha = 0.6, color = NA) +
  
  geom_boxplot(width = 0.15, outlier.shape = NA, alpha = 0.8) +
  
  #geom_jitter(width = 0.1, size = 1.2, alpha = 0.5) +
  geom_signif(
    comparisons = list(c("onestep", "twostep")),
    test = "wilcox.test",
    map_signif_level = TRUE,
    y_position = max(sum_dat$Value) * 1.05,
    tip_length = 0.01,
    textsize = 5
  )+
  scale_fill_manual(values = c( group_valuess[[1]],group_valuess[[2]])) +
  theme_yzc
dev.off()








```
